Effects of Different Pain Relievers and Dosages on Cognitive Retention

### 2025-04-05 paper/paper.qmd
Effect Size and Power Analysis

## Effect Size

@@

Instead, we refer to Cohen’s \(f\) as a more appropriate measure of effect size in ANOVA. Cohen’s \(f\) is based on the proportion of variance explained and provides context for the magnitude of the observed effects. For this study, we assume a medium effect size, corresponding to Cohen’s $f \approx 0.25$, based on conventional benchmarks from lecture materials.

Even though the statistical test did not yield significant results. This suggests a small-medium effect size, indicating that the type of drug and dosage level may still have had a meaningful impact on memory task performance.

@@

>>>>

## Effect Size and Power Analysis

@@

Instead, we refer to Cohen’s \(f\) as a more appropriate measure of effect size in ANOVA. Cohen’s \(f\) is based on the proportion of variance explained and provides context for the magnitude of the observed effects. [[[SHOULD WE STILL INCLUDE THIS?: For this study, we assume a medium effect size, corresponding to Cohen’s $f \approx 0.25$, based on conventional benchmarks from lecture materials. Even though the statistical test did not yield significant results. This suggests a small-medium effect size, indicating that the type of drug and dosage level may still have had a meaningful impact on memory task performance.]]]

We did a post-hoc power analysis using @WebPower. Based on the overall $\eta^2$ from the ANOVA model, we calculated a Cohen's f value of 0.204. This suggests a small-medium effect size, which means that the type of drug and dosage level may still have had a meaningful impact on memory task performance.

The estimated power of the study was around 59.6%. This means that our study was moderately able to detect an effect with the size we have, if there was an effect. Despite being less than the ideal 80% power value, this suggests that the study is still able to detect large & meaningful differences -- just not smaller ones.

@@

```{r}
# EFFECT SIZE AND POWER ANALYSIS
eta_sq <- eta_squared(anova_model, partial = FALSE)

etasquare <- sum(eta_sq$Eta2)  # sum of all eta^2 values
cohens_f <- sqrt(etasquare / (1 - etasquare))

power_result <- wp.kanova(
  n = 240,
  ndf = 7, # because 8 - 1 = 7
  f = cohens_f,
  ng = 8, # because 8 treatment groups
  alpha = 0.05
)
